Do Machines Think Equally About Music?

### 2025-04-28 paper/paper.qmd
add toc to paper; okay now i'm done

---
title: "Do Machines Think Equally About Music?"
subtitle: "Testing an SVM's Classification Ability on Machine-Generated Music"
author: 
    - Luca Carnegie
thanks: "Code, workflow and prompts are available at: https://github.com/lcarnegie/song-classification. Many thanks to Periklis Andritsos for your supervision, guidance and excellent conversation."
date: today
date-format: long
abstract: "This paper demonstrates a systematic workflow for measuring how well a Music Generation Model, such as Meta's MuseGen, has learned the patterns of a particular genre or style of music. First, several classifier models are trained in Python, including an SVM, K-nearest-neighbors, Random Forest, and XGBoost, with the SVM being the most balanced at classification across genres. Then, the smallest version of Facebook's MuseGen model is systematically prompted through the MusicGPT interface to generate music 'quintessentially' describing one of ten genres of music. Finally, the features of this music are extracted using Python and then individually inputted into the SVM model, with the trained model in this workflow setup unable to classify any of the generated music clips. This paper highlights the need for further work on how machine learning algorithms embed understanding about musical features like timbre and how machine-generated music can impact existing music information retrieval systems, providing directions for future work in this important area. These points are particularly pertinent in light of extensive developments in machine-generated music."
format: pdf
number-sections: true
bibliography: references.bib
---

>>>>

---
title: "Do Machines Think Equally About Music?"
subtitle: "Testing an SVM's Classification Ability on Machine-Generated Music"
author: 
    - Luca Carnegie
thanks: "Code, workflow and prompts are available at: https://github.com/lcarnegie/song-classification. Many thanks to Periklis Andritsos for your supervision, guidance and excellent conversation."
date: today
date-format: long
abstract: "This paper demonstrates a systematic workflow for measuring how well a Music Generation Model, such as Meta's MuseGen, has learned the patterns of a particular genre or style of music. First, several classifier models are trained in Python, including an SVM, K-nearest-neighbors, Random Forest, and XGBoost, with the SVM being the most balanced at classification across genres. Then, the smallest version of Facebook's MuseGen model is systematically prompted through the MusicGPT interface to generate music 'quintessentially' describing one of ten genres of music. Finally, the features of this music are extracted using Python and then individually inputted into the SVM model, with the trained model in this workflow setup unable to classify any of the generated music clips. This paper highlights the need for further work on how machine learning algorithms embed understanding about musical features like timbre and how machine-generated music can impact existing music information retrieval systems, providing directions for future work in this important area. These points are particularly pertinent in light of extensive developments in machine-generated music."
toc: true
format: pdf
number-sections: true
bibliography: references.bib
---
